### 2017-03-31 egpaper_for_review.tex
Initial commit.

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi

\usepackage{color}

\newcommand{\matt}[1]{ \color{red} Matt: #1  \color{black}}
\newcommand{\aneesh}[1]{ \color{blue} Aneesh: #1  \color{black}}
\newcommand{\burak}[1]{ \color{green} Burak: #1  \color{black}}
\newcommand{\emmett}[1]{\color{violet} Emmett: #1  \color{black}}
\begin{document}

%%%%%%%%% TITLE
\title{High Speed Object Tracking using Multiple Kernelized Correlation Filters}

\author{Burak Uzkent\\
Rochester Institute of Technology\\
{\tt\small bxu2522@@rit.edu}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
YoungWoo Seo\\
Affiliation\\
{\tt\small yseo@autelrobotics.edu}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}

\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
The tracking-by-detection methods have recently emerged as robust, powerful, and highly discriminative trackers replacing the preceding generative trackers that focused on only target modeling. Some of the recent tracking-by-detection algorithms are Tracking-Learning-Detection \cite{kalal2012tracking}, Multiple Instance Learning \cite{babenko2009visual}, Correlation Filter based trackers \cite{bolme2010visual,henriques2015high}, and Struck \cite{hare2012efficient}. These tracking-by-detection methods train an object-specific classifier on the fly with the focus of discriminating the target from the background. Learning an object-specific classifier in an online manner can better handle challenging scenarios with the cost of increased complexity. In order to train a classifier, one need positive and negative samples. Randomly collecting negative samples from the background area can be one straightforward way of building training data to train the classifier. However, there can be two issues in this framework. First, the randomly collected samples can have large correlation resulting in a weak classifier. Also, such framework is not feasible for real-time tracking as the number of samples gradually increases. Most of the tracking-by-detection algorithms focus on efficient incremental update of a classifier together with more useful background sampling.

Frequency domain correlation filter based trackers in the tracking-by-detection paradigm has recently seen growing popularity due to its superior computation and robustness to geometric and photometric variations. The first correlation filter tracker proposed by \cite{bolme2010visual} focused on learning and detection in the frequency domain. The idea was to minimize the sum of squared error in the frequency domain. Hence, the convolution in the original domain is avoided, instead, element-wise multiplication is needed in the frequency domain. This way, it achieved promising performance at 700 Hz. On the other hand, the MOSSE tracker failed in cases with rapidly changing background as it learns the filter using a single template and accommodates single channel features. Later on, the MOSSE tracker was improved by using the circularly shifted background patches in an efficient way by utilizing the theorem that Discrete Fourier Transform of a circulant matrix gives us diagonalized circulant matrix \cite{henriques2012exploiting,henriques2015high}. In the same work, the authors integrated kernelization concept into the detection and learning steps to improve tracking accuracy. \cite{galoogahi2013multi,henriques2015high} extended the correlation filter trackers with single channel features to multi-channel features. They employed the color and shape features in a simple framework by summing the correlation of the test and learned feature channels elements. There a few major drawbacks of the correlation filter trackers although they outperformed the other discriminative trackers with superior computation. First, due to the nature of correlation filters it is limited to a region of interest which is typically 2-3 times larger than the target. Considering a larger ROI not only increases the computational burden but also leads to further spatial information loss with the fixed template size. For this reason, its performance degrades in fast motion cases as the ROI may not contain the target. Second, the correlation filter lacks scale adaptation ability since it only estimates the translation of the target. Later studies focused on improving the latter issue by proposing scale candidates where each scale is assigned a confidence by performing detection on the shrinked or enlarged the ROI \cite{li2014scale,tang2015multi,bibi2015multi,ma2015long}. Another group of studies employ the part-based correlation filter trackers to better adapt to the scale changes \cite{liu2015real,akin2016deformable}. The sensitivity to fast motion is generally handled in the long-term target tracking framework where track reinilization module is triggered when the target is lost \cite{ma2015long,de2015board,li2016monocular}.

CONTRIBUTION

%--------------
%\section*{Acknowledgements}
%put stuff here for the accepted , but not the ICCV version

{\small
\bibliographystyle{ieee}
\bibliography{egbib}
}

\end{document}