### 2017-04-28 draft.tex
Revised the introduction and todo list.

\documentclass[10pt,twocolumn,letterpaper]{article}
\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools} % mathtools builds on and extends amsmath package
\usepackage{algorithm}		% http://ctan.org/pkg/algorithms
\usepackage{algpseudocode}	% http://ctan.org/pkg/algorithmicx% Include other packages here, before hyperref.
\usepackage{comment, url}

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}
% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi

\usepackage{color}

\newcommand{\matt}[1]{ \color{red} Matt: #1  \color{black}}
\newcommand{\aneesh}[1]{ \color{blue} Aneesh: #1  \color{black}}
\newcommand{\burak}[1]{ \color{green} Burak: #1  \color{black}}
\newcommand{\emmett}[1]{\color{violet} Emmett: #1  \color{black}}
\usepackage{forloop}
\newcounter{ct}
\newcommand{\markdent}[1]{\forloop{ct}{0}{\value{ct} < #1}{\hspace{\algorithmicindent}}}
\newcommand{\markcomment}[1]{\Statex\markdent{#1}}
\begin{document}

%%%%%%%%% TITLE
\title{ {\it E}nKCF: Ensemble of Kernelized Correlation Filters for Object Tracking in High Speed}

\author{Burak Uzkent\\
Rochester Institute of Technology\\
{\tt\small bxu2522@@rit.edu}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
YoungWoo Seo\\
Independent Robotics Research\\
{\tt\small youngwoo.blank.seo@gmail.com}
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
Computer vision technologies are very attractive for practical
applications for embedded systems in that most of the embedded systems
already includes image acquisition pipeline and a tremendous amount of
progresses on research has been made for many areas. However, to
successfully deploy a computer vision algorithm on any existing
embedded systems, a vision algorithm needs to satisfy some criteria:
minimal, manual intervention after deployment and small footprint of
computing resource consumption and on executable code, with the
assumption of reasonably good performance. To this end, in this paper,
we propose an ensemble of the kernelized correlation filters (KCF), we
call {\it E}nKCF, for a single-target object tracking. In particular,
we developed a committee of KCFs to specifically address the
scale-change and dynamic maneuver of the target over frames. In
addition, we devised a Bayes filter for a smooth transition between
individual KCFs' executions. To minimize the effect of drifting, we
also developed another algorithm that detects target-lost, uses an
existing object proposal method (e.g., ``Edge Box'') to re-detect the
lost target, and re-initializes the previous tracking for a long-term
tracking. To verify the usefulness of the proposed method, we compared
the performance of ours to those of the existing online tracking
methods. Experimental results showed that the performance of ours are
72.53\% for precision on 20 pixels and 52.90\% for success rate for
OTB100 data, and 55.71\% and 41.71\% for UAV123 data. These results
confirmed that our method outperformed the existing ones over 5\% on
precision on 20 pixels and 10-20\% on AUC on average. Our
implementation ran at 340 fps for OTB100 and at 416 fps for UAV123
data that is faster than DCF (292 fps) for OTB100 and KCF (292 fps),
DCF (457 fps), and STC (350 fps) for UAV123.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
\begin{itemize}
\item YoungWoo: Clean up the introduction
\item Burak: Clean up the technical section: EnKCF, KCF, and Particle Filter
\item Burak: You'll have a conclusion whether we'll have the re-detection part by Friday night.
\item By Monday, we'll have a draft ready to submit.
\end{itemize}

A recent proliferation of air/ground/water unmanned vehicle
technologies has ever increased interests on deployment of intelligent
algorithms on embedded or mobile platforms. Among those technologies,
computer vision algorithms are getting more attentions primarily
because those mobile platforms have limited payload to carry other
active sensors than a lightweight, monocular camera and, without
increasing much cost, installation of a monocular camera even stereo
rigs would increase the usefulness of a mobile platform. For example,
instead of having just bare flight function with video recording, an
unmanned air vehicle (UAV) equipped with object or feature following
would make it more useful in the application of
monitoring/surveillance/surveying on private
properties/wild-life/crop, video recording on sports/movies/events,
many others.

In this paper, we propose a single-target tracking that aims at
running on any embedded or mobile platforms that doesn't require an
offline training, and can run on an embedded system. In particular, we
would like to make our algorithm 1) learn the appearance model of a
target on the fly and 2) run as fast on a typical desktop as up to
300Hz so that it could run up to 30Hz when it is deployed on a
low-end, embedded system. With these features, our tracking algorithm
is ready to track an object of interest when it is deployed on an
embedded system. To this end, an online object tracking is more
appropriate.

{\it We may need to change this introductory sentence based on the
  results.}

One of the dominant frameworks for online object tracking is the
correlation filter that essentially solves a single-target tracking
problem as a regression problem in the frequency domain. This
framework assumes that the target location is given at the very first
image frame like other online tracking algorithms
\cite{smeulders2014survey}. Given this positive example for the
regression problem, a set of negative examples is collected around the
initial, target bounding box and represented as a circulant matrix
\cite{henriques2015high}. One can optimally solve this regression
problem using a ridge regression in a closed form. However, such a
solution involves in expensive matrix operations
$\mathcal{O}(n^{3})$. The correlation filter provides one with a much
cheaper solution, $\mathcal{O}(n\log n)$ over element-wise
multiplication in a frequecny domain
\cite{bolme2010visual,henriques2015high}. Thank to this reformulation,
a complete pipeline of an object tracker based on the correlation
filter can be easily implemented and run very efficiently in an online
fashion. In fact, a variant of the correlation filter, kernelized
correlation filter with multi-channel features, outperformed other
state-of-the-art tracking algorithms in VOT15 challenge, in terms of
run-time and tracking accuracy. However, the vanilla form of such an
online tracker is prone to drift, and fails to track a target over a
long period of time primarily \cite{henriques2015high} because of the
dilemma of stability-plasticity in updating appearance model, where
the appearance model will be overfitted to only the images used to
train and a compromise on the frequency of updating the model is
carefully implemented \cite{santner2010prost}. For example, a naive
way of handling a scale change of a target could reduce the run-time
performance from 300 to 100 fps. A seemingly obvious way of addressing
the scale change is to scan the region of interest (ROI) with
templates in different dimensions based on pre-defined scale ratios to
find the target
\cite{henriques2015high,tang2015multi,ma2015long,bibi2015multi,li2014scale}. However,
because of computational nature of correlation filter, this approach
would just increase the computational complexity of tracking by
running detection with multiple possibilities.

Another way of handling the scale change for the correlation filter
based approach is to find a correct scale at a location where a target
highly likely appears \cite{zhang2014fast}. In particular, they uses
the MOSSE tracker to estimate a target's translation. The scale is
updated in a naive way where the confidence map is used to determine
the scale change between successive frames.

For instance, their method assumes similar scale in between two
consecutive frames given similar confidences. With this simple method,
we do not need to run detection on different ROIs at a given frame and
perform tracking at $\geq300$ fps. In this study, we propose the use
of Ensemble KCFs in an intuitive way to perform robust scale-adaptive
tracking at more than $300$ fps. Our approach is inspired by
\cite{ma2015long} where two KCFs are employed as translation and scale
filters. It first estimates translation using the translation filter
learned on \textit{target}$+$\textit{background} area. Then, the scale
filter learned on the \textit{target} area is used to estimate new
scale of the target at the estimated position. Application of
translation and scale filter at each frame reduces run-time
performance ($\leq50 fps$) as we need to run detection on each scale
candidate and perform training on both filters. Similarly, we learn
three different correlation filters specializing on different aspects
of tracking and addresses the weakness of each other, resulting in
more robust tracking. We name these filters as
\textit{target}+\textit{small background} translation filter
($R_{t}^{S}$), \textit{target-only} scale filter ($R_{s}$) and
\textit{target}+\textit{large background} translation filter
($R_{t}^{L}$). Unlike \cite{ma2015long}, by running a single filter at
each frame, we can achieve the targeted operation frame rate. The
fig.~\ref{fig:Filters} shows the size of the ROIs associated with
$R_{t}^{S}$, $R_{t}^{L}$, and $R_{s}$ in addition to the hanning
windows assigned and desired Gaussian responses assigned to each one.

The recent progress on development of convolutional neural networks
(CNN) is astonishing as since the AlexNet \cite{krizhevsky12} broke
the record for the ImageNet competition, at every year, new and better
results are reported. Moreover, there have been many attempts to
deploy CNN on the mobile devices
\cite{wu2016quantized,giusti2015}. But the inherent limitation of such
a supervised learning based object tracking is that an either online
\cite{nam2016} or offline \cite{held2016} training tracker has to be
prior to deploy the algorithm.

The contributions of this study are as follows.
\begin{itemize}

\item \textbf{Novel Single-Target Tracking in High-Speed for Long-Term
  Tracking} To be practical, the execution time of an object tracking
  is important. We proposed a new object tracking algorithm that is a
  collection of kernelized correlation filters (KCF). Each of KCFs is
  designed to address two challenges of object tracking: scale and
  translation. We also incorporate a Bayes filter to prevent from
  being drift.

\item \textbf{Re-Detection} A target loss is inevitable in object
  tracking. To cope with this, we developed a new target, re-detection
  that reliably detects when to lose the target and effectively
  re-detect the lost target. {\it How we can show how good our
    re-detection method is?}

\begin{comment}
\item We build a highly efficient ($\geq300$ fps) scale adaptive
  multiple kernelized correlation filter based tracker that
  outperforms the original KCF implementation with fixed scale
  framework both in terms of accuracy and performance. The studies
  following the KCF improved the fixed scale framework by running
  detection on a number of candidate ROIs to figure out the new scale
  of the object after estimating the translation of the object. This
  approach adds additional complexity to the original KCF
  implementation and drags down the run-time performance from $300$
  fps to less than $100$ fps.

\item We integrate the Particle Filter into the Multiple KCFs tracking
  as an additional filter that can avoid the drift due to one of the
  correlation filters we employ in our framework.

\item We propose a target re-detection module that can minimize the
  target loss due to the scale filter that learns the object model
  using only-object area. Also, the target re-detection step is
  required to handle severe occlusions, pose variations, illumination
  changes and fast motion.

\item Since our visual tracker mostly focuses on tracking objects from
  aerial moving platforms, we design a new robust target-to-camera
  distance estimation method. This way, a safe distance between the
  target and the camera-platform can be preserved.
\end{comment}

\end{itemize}

%---------------------------------------------------------------------- 
\section{{\it E}nKCF: Ensemble of Kernelized Correlation Filters}
%----------------------------------------------------------------------

The workflow of our MKCF based long-term target following method can
be visualized in fig.~\ref{Workflow_figure}. Summarize what we discuss
in the next sections.
\begin{figure*}[!t]
\includegraphics[width=\textwidth]{figures/Workflow_MKCF+PF.pdf}
\caption{The workflow of our Multiple KCF and Particle Filter based tracking method without the track re-detection module.}
\label{Workflow_figure}
\end{figure*}

The KCF with multi-channel features outperform the other
state-of-the-art object traking algorithms in VOT15 challenge in terms of run-time
performance and tracking accuracy. However, it lacks scale adaptation
module or the naive scale adaptation method added into the KCF reduces
the frames per second from $\geq300$ fps to less than $100$ fps. One
straightforward method is to run detection on ROIs with different
sizes determined by pre-defined scale ratios
\cite{henriques2015high,tang2015multi,ma2015long,bibi2015multi,li2014scale}. All
these method increases the computational complexity of tracking by
running detection on a number of ROIs. Another scale update method in
Correlation Filter framework was proposed by \cite{zhang2014fast}. It
uses the MOSSE tracker to estimate translation of an object. The scale
is updated in a naive way where the confidence map is used to
determine the scale change between successive frames. For instance,
their method assumes similar scale in between two consecutive frames
given similar confidences. With this simple method, we do not need to
run detection on different ROIs at a given frame and perform tracking
at $\geq300$ fps. In this study, we propose the use of Ensemble KCFs
in an intuitive way to perform robust scale-adaptive tracking at more
than $300$ fps. Our approach is inspired by \cite{ma2015long} where
two KCFs are employed as translation and scale filters. It first
estimates translation using the translation filter learned on
\textit{target}$+$\textit{background} area. Then, the scale filter
learned on the \textit{target} area is used to estimate new scale of
the target at the estimated position. Application of translation and
scale filter at each frame reduces run-time performance ($\leq50 fps$)
as we need to run detection on each scale candidate and perform training on both filters. Similarly,
we learn three different correlation filters specializing on different
aspects of tracking and addresses the weakness of each other,
resulting in more robust tracking. We name these filters as
\textit{target}+\textit{small background} translation filter
($R_{t}^{S}$), \textit{target-only} scale filter ($R_{s}$) and
\textit{target}+\textit{large background} translation filter
($R_{t}^{L}$). Unlike \cite{ma2015long}, by running a single filter at
each frame, we can achieve the targeted operation frame rate. The
fig.~\ref{fig:Filters} shows the size of the ROIs associated with
$R_{t}^{S}$, $R_{t}^{L}$, and $R_{s}$ in addition to the hanning
windows assigned and desired Gaussian responses assigned to each one.

\begin{figure}[!t]
\includegraphics[width=0.5\textwidth]{figures/Filters_Details.pdf}
\caption{The three filters denoted as \textit{target+small background} translation filter,
  \textit{target-only} scale filter and \textit{target+large background} translation filters are
  shown. Also, the hanning window and desired Gaussian response for
  each filter is displayed.}
\label{fig:Filters}
\end{figure}

Our MKCF based tracking framework employs different capabilities and
tasks to each correlation filter learned temporally. In
\cite{ma2015long, danelljan2014accurate}, a scale filter is learned on
\textit{target-only} area whereas \cite{henriques2015high,
  li2014scale, bibi2015multi, tang2015multi} uses a correlation filter
learned on \textit{target+background} area to find the optimal scale
in scale space. To learn the scale filter, $R_{s}$, we follow a
similar approach to \cite{ma2015long}. We believe that translation
filter needs to include background information to better discriminate
target from background, however, the goal of the scale filter is to
estimate the right size of the target, not to discriminate it from
background. Also, by using a smaller ROI we can reduce the template size
 to learn the scale filter, resulting in higher speed scale
estimation module. Finally, the scale filter help us on the track
re-initialization module which includes running an object proposals
algorithm to boxes bounding the objects in the scene. By using the
scale filter in re-initialization module, we can reduce the
element-wise operations compared to the translation filters. Also, we can 
avoid false target re-initialization in the case of rapid background change as $R_{t}^{s}$ and
$R_{t}^{L}$ learn background in addition to target. The \textit{target+small background} area translation filter, $R_{t}^{S}$,
is assigned a larger template size than $R_{s}$ and smaller than
$R_{t}^{L}$. It is employed to estimate translation and scale similar
to \cite{henriques2015high}. Finally, the \textit{target+large
  background} area translation filter, $R_{t}^{L}$ , estimates only
the translation of a target. We propose scale estimation at every $n$ number of frames using the scale filter. In our experiments on $OTB100$ and $UAV123$,
we observed that $n$ = 5 gives the best results in terms of precision and success overlap rates. The \textit{target+large background} translation filter is run
in the next frame after the scale filter as we want to avoid drifts. Considering large ROI for $R_{t}^{L}$ can degrade its performance due to low resolution template. For this reason, we employ \textit{Color} and \textit{HoG} features for $R_{t}^{L}$ correlation filter. Finally, $R_{t}^{S}$ is applied after $R_{t}^{L}$ to avoid any target loss due to low resolution template used by $R_{t}^{L}$. Unlike $R_{t}^{L}$, we only employ the \textit{HoG} features for the $R_{t}^{S}$ correlation filter. Fig.~\ref{Workflow_figure} displays the proposed tracking framework using Ensemble of KCFs. Also, the pseudo-code of the proposed $EnKCF$ is shown in alg.~\ref{alg:MKCF}.

\begin{algorithm}
	\caption{The EnKCF Tracking Algorithm.}\label{alg:MKCF}
	\begin{algorithmic}[1]
	\State \textbf{Input} : Initial bounding box $x_{0}$, frame counter $fc$, scale filter frequency $n = 5$,
	\State \textbf{Output} : 
				\If{$fc\:\%\:n=0$} \Comment{Condition 1}
						\State Estimated Target State $x_{t} = (x_{t-1},y_{t-1},s_{t})$,
						Scale filter (\textit{target-only}) model $R_{s}$.
			     \EndIf
				\If{$fc\:\%\:n=1\:\:or\:\:fc\:\%\:n=2$}\Comment{Condition 2}
						\State Estimated Target State $x_{t} = (x_{t},y_{t},s_{t_1})$,
						Large Area Translation Filter model $R_{t}^{L}$.
				\EndIf
				\If{$fc\:\%\:n=1\:\:or\:\:fc\:\%\:n=4$}\Comment{Condition 3}
						\State Estimated Target State $x_{t} = (x_{t},y_{t},s_{t})$,
						Small Area Translation Filter model $R_{t}^{S}$.
				\EndIf
	\Procedure{track}{$x_{t-1},y_{t-1},s_{t-1}$} 
				\State // Translation Estimation
				\State Transit Particle Filter to the frame $t$ and compute the mean of prior pdf $x_{t} = (x_{t},y_{t},s_{t-1})$;
				\State // Translation Estimation
				\State Crop the ROI for the $R_{t}^{L}$, or $R_{t}^{S}$ given $x_{t}$
				and estimate translation ($x_{t}$,$y_{t}$) using $R_{t}^{L}$ (Condition 2) or $R_{t}^{S}$ (Condition 3),
				\State Skip translation estimation for $R_{s}$ (Condition 1);
				\State // Scale Estimation
			    \State Crop the ROI for the $R_{s}$ and estimate scale, $s_{t}$, using $R_{s}$ (Condition 1), 
		         \State Skip it for $R_{t}^{L}$ (Condition 1),
				\State Scale pool for $R_{s}$ : $\lbrace1.05,1.0,1/1.05\rbrace$;
				\State // Update Translation
				\State Perform Importance Re-sampling for Particle Filter and compute the mean of posterior pdf $x_{t} = (x_{t},y_{t},s_{t})$;
			    \State // Model Update
				\If{$PSR(y_{R_{s}}) \geq T_{R_{s}}$}
				\State Update $R_{s}$ (Condition 1);
				\EndIf							 
				\If{$PSR(y_{R_{t}^{L}}) \geq T_{R_{t}^{L}}$} 
				\State Update $R_{t}^{L}$ (Condition 2);
				\EndIf	
			     \If{$PSR(y_{R_{t}^{S}}) \geq T_{R_{t}^{S}}$}
				\State Update $R_{t}^{S}$ (Condition 3);
			     \EndIf		
	\EndProcedure	
	\end{algorithmic}
\end{algorithm}

The Kernelized Correlation Filter tracker has recently been
increasingly popular due to its operation at hundreds of frames per
second with state-of-the-art tracking capabilities in challenging
cases. Its computational efficiency arises from its use of the
discrete fourier transform of the circulant matrix and frequency
domain element-wise operations knows hadamard product and
division. The first example of frequency domain trackers is the Linear
Correlation Filter tracker known as MOSSE tracker. It minimizes the
ridge regression function in the frequency domain using a single
template with continous desired gaussian response. The KCF tracker, on
the other hand, minimizes the regularized ridge regression function
shown below.
\begin{equation}
E(h) = \frac{1}{2}||y-\sum_{c=1}^{C}g*x_{c}||^{2} + \frac{\lambda}{2}\sum_{c=1}^{C}||h_{c}||^{2}
\label{eq:Closedform_RidgeReg}
\end{equation}
where $y$ represents the desired continous response whereas $h$ and
$x_{c}$ represents the learned correlation filter and training
template for the given channel. The $c$ parameter included in
\cite{henriques2015high,galoogahi2013multi} makes it possible to
integrate multi-channel features such as HoG and colour into the ridge
regression function. The closed-form solution for the
eq.~\ref{eq:Closedform_RidgeReg} can be obtained by setting the
derivative of $E$ w.r.t $w$ to $0$. The solution in the primal domain
can be formulated as
\begin{equation}
w = (X^{T}X+\lambda)^{-1}y
\label{eq:SpatialSolution}
\end{equation}
where $X$ and $\lambda$ represent the training samples and
regularization term. The same cost function in the Fourier domain can
be represented as
\begin{equation}
w = (X^{H}X+\lambda I)^{-1}X^{T}y
\label{eq:FourierSolution}
\end{equation}
More information on the spatial and fourier domain solutions can be
found in \cite{henriques2015high}. The MOSSE tracker does not make use
of $\lambda$ and only one training sample with desired response is
used to update $w$. This framework does not include enough background
information into the training framework with a single template. The
application of the circulant matrix theorem into the
eq.~\ref{eq:Closedform_RidgeReg} makes it possible to include many
background patches at a similar computational complexity. A circulant
matrix $C$ includes the circularly shifted patches of the positive
training sample $x$ by the cylic shift operator $P$. By applying
shifting operation to the base sample, we can generate the circulant
matrix as
\begin{equation}
C = Px.
\label{eq:CirculantMatrixGeneration}
\end{equation}
The circulant matrix of a base sample $x$ can be interpreted as the
rows of a training matrix $X$ where each row represents features of a
training sample. Mathematically, this can be written as
\begin{equation}
X = C(x)
\label{eq:CIrculantMatrixTrainingData}
\end{equation}
In this form, eq.~\ref{eq:SpatialSolution}
and~\ref{eq:FourierSolution} could prohibit us from implementing a
high speed object tracking as we need to perform large number of
element-wise division and multiplication operations. However, we know
from \cite{gray2006toeplitz} that all circulant matrices are
represented diagonally by the Fourier Transform regardless of the base
sample $x$ as shown below.
\begin{equation}
X = Fdiag(\hat{x})F^{H}.
\label{eq:CirculantMatrixDFT}
\end{equation}
where $F$ denotes a constant Fourier Transform matrix and $x$ is the
Discrete Fourier Transform of the base sample $x$. To simplify the
cost function formulation in eq.~\ref{eq:FourierSolution} we multiply
$X$ in eq.~\ref{eq:CirculantMatrixDFT} with $X^{H}$ yielding
\begin{equation}
X^{H}X = Fdiag(\hat{x}^{*}\odot \hat{x})F^{H}.
\label{eq:SimplificationX} 
\end{equation}
Finally, the eq.~\ref{eq:SimplificationX} can used to formulate the
solution vector in Fourier domain $\hat{w}$ as
\begin{equation}
\hat{w} = \dfrac{\hat{x}^{*}*\hat{y}}{\hat{x}^{*}*\hat{x}+\lambda}.
\label{eq:DiagonalizedPrimalSolution}
\end{equation}
For detailed documentation of the circulant matrix theorem based
frequency domain solution can be found in
\cite{henriques2012exploiting,henriques2015high}.

The above primal frequency domain solution can be called as Linear
Correlation Filter which improves the MOSSE tracker by incorporating
cyclic shifts and regularizer. To further improve the robustness to
geometric and photographic variations, one can exploit non-linear
regression function in the Correlation Filter framework
\cite{henriques2015high}. The solution to the kernelized ridge
regression function is shown below.
\begin{equation}
\alpha = y(K+\lambda I)^{-1}
\end{equation}
wjere $K$ and $\alpha$ represent the kernel matrix and corresponding
dual space solution. \cite{henriques2015high} states that kernel
matrices is circulant for datasets of circular cylic satisfying the
following theorem.
\begin{equation}
k(x,x^{'}) = k(Mx,Mx^{'})
\label{eq:KernelCirculantTheorem}
\end{equation}
where $M$ represents the permutation matrix. Some kernels satisfying
the eq.~\ref{eq:KernelCirculantTheorem} are \textit{Gaussian},
\textit{Polynomial}, \textit{Intersection} and \textit{Hellinger}
kernels. Similar to the diagonalization in the linear ridge regression
solution in eq.~\ref{eq:DiagonalizedPrimalSolution}, the kernelized
ridge regression can be made diagonal using the same circulant matrix
theorem.  The diagonalized Fourier domain dual form solution can be
expressed as
\begin{equation}
\hat{\alpha} = \hat{y}(\hat{k}^{xx}+\lambda)^{-1}
\label{eq:FourierDualDomainSolution}.
\end{equation}
where $\hat{k}^{xx}$ represents the first row of the Kernel matrix $K$
known as \textit{gram matrix}. In this study, we will only focus on
application of the Gaussian Kernel to the Correlation Filters. We
refer the readers to \cite{henriques2015high} for the detailed
documentation of the application of other kernels to the Correlation
Filters. For single channel features, the Gaussian kernelization is
expressed as
\begin{equation}
k^{xx^{'}} = exp(-\dfrac{1}{\alpha^{2}}(||x||^{2}+||x^{'}||^{2}-2F^{-1}(\hat{x}^{*}\odot \hat{x}^{'})))
\label{eq:GaussianCorrelationSingleChannel}
\end{equation}
The first correlation filter based trackers used grayscale feature to
learn the solution vector $w$, however, the multi-channel features
such as HoG and Color were later exploited to improve tracking
accuracy
\cite{henriques2015high,galoogahi2013multi,tang2015multi,ma2015long,bibi2015multi}. The
multi-channel feature integration into the Gaussian Kernelization
function in eq.~\ref{eq:GaussianCorrelationSingleChannel} is achieved
in a very straight-forward way by summing the correlation result in
all the channels as
\begin{equation}
k^{xx^{'}} = exp(-\dfrac{1}{\alpha^{2}}(||x||^{2}+||x^{'}||^{2}-2F^{-1}(\sum^{C}_{c}\hat{x}_{c}^{*}\odot \hat{x}_{c}^{'}))).
\label{eq:GaussianCorrelationSingleChannel}
\end{equation}
Such non-linearization process does not increase the computational
complexity of the linear multi-channel correlation filter dramatically
as we only need to sum over the $n$ dimensional feature
channels. Training shown in eq:~\ref{eq:FourierDualDomainSolution}
gives us $\hat{a}$ learned in time step $t$. We can accumulate
$\hat{a}$ over time to integrate more temporal information. This can
be expressed as
\begin{equation}
\hat{a}_{t} = (1-\beta)\hat{a}_{t-1} + \beta\hat{a}_{t}. 
\end{equation}

Finally, detection step in multi-channel KCF framework is performed as
\begin{equation}
r(z) = F^{-1}(\hat{k}^{xz} \odot \hat{\alpha})
\end{equation}
where $r$ denotes the correlation response at all cylic shifts of the
first row of the kernel matrix $K$. The peak point of the response
function gives us the estimated translation of the object from time
step $t$-$1$ to $t$.

\section{Particle Filter}
\label{sc:PF}
The proposed MKCF algorithm can cause undesired drift and noise in
tracking due to independent application of the correlation filters. To
mitigate the drift effect and reduce the noise, a Bayes Filter
representing the evolution of target's motion can be added into the
MKCF tracking framework. In this study, we consider a variant of a
Bayes Filter, \textit{Particle Filter}. The Particle Filter is a
Sequential Monte Carlo method that can approximate posterior
probability density functions (pdf) of a target motion. The
approximation becomes optimal with the infinite number of particles,
however, its run-time complexity grows exponentially with the number
of elements in the state space matrix, $X_{t}$. As the focus of this
study is to design a high speed tracker, we keep the number of state
space elements low and implement computationally cheap \textit{Weight
  Function} and \textit{Importance-Resampling} modules. We represent
the state space matrix as $X_{t} = \lbrace x,y,V_{x},V_{y}\rbrace$
where $x$ and $y$ are the centroid of the target and $V$ represents
the velocity components. We transit the particles with the well-known
first order Constant Velocity model. The centroid components are
applied Gaussian Noise whereas the velocity components are assigned
uniform noise representative of most of the objects. The observation
likelihood is modeled based on the confidence map of
$R_{t}^{L}$,$R_{t}^{S}$, and $R_{s}$. It can also be modeled based on
the euclidean distance between particle's centroid to the peak of the
confidence maps, however, we believe that the former approach could
result in further drifts in the case of multi-modal distributions. The
weights for the particles are computed with the bilinear interpolation
method.
\begin{equation}
	w_{p_{t}}(x_{t},y_{t}) = \dfrac{\splitfrac{y_{R}(x_{t}-1,y_{t}-1)+y_{R}(x_{t}+1,y_{t}+1)+}{y_{R}(x_{t}+1,y_{t}-1)+y_{R}(x_{t}-1,y_{t}+1)}}{4}
\end{equation}
where $w_{p}$ denotes the weight of the particle. The importance
re-sampling module is triggered when the number of effective particles
is smaller than a pre-defined threshold [Cite - Effective Number
  Samples Metric] to avoid variance getting too small. Re-sampling is
implemented with highly efficient low-variance ratio method
[cite]. Finally, the expected mean of the approximated posterior pdf
is computed as
\begin{equation}
	\hat{X}_{t} = \sum_{p=1}^{P}w_{p_{t}} X_{t}
\end{equation}

The particles are assigned equal weight ($\dfrac{1}{P}$) after the
importance re-sampling step. It should be emphasized that we skip
importance re-sampling step in the scale filter operation as the scale
filter contains \textit{target-only} area. In this case, the particles
outside of the scale filter ROI do not have weight correspondence in
the confidence map. Also, it is more likely to get noisy translation
estimation from the scale filter.

WE SHOULD ASSIGN THE PRIOR WEIGHTS TO EXPECTED MEAN ESTIMATION OF THE
POSTERIORS WHEN THE RE-SAMPLING IS SKIPPED.
\section{Track Re-initilization via Re-detection}
\label{sc:Re-initialization}

%\section{Geometry Estimation between Target and Drone}
%\label{sc:Geometry}

%----------------------------------------------------------------------
\section{Experiments} \label{sc:Experiments}
%---------------------------------------------------------------------- 
To verify the usefulness of our algorithm, we ran our algorithm using
two publicly available datasets:
%VOT2015 \footnote{\url{http://www.votchallenge.net/}},
OTB100 \footnote{\url{http://cvlab.hanyang.ac.kr/tracker_benchmark/benchmark_v10.html}},
UAV123 \footnote{\url{https://ivul.kaust.edu.sa/Pages/Dataset-UAV123.aspx}}\cite{mueller2016uav123}. OTB100 dataset contains video sequences for 100 objects whereas UAV123 dataset contains aerial footage of 123 objects. The proposed $EnKCF$ method can better handle slow motions. Also, the UAVs use mobiles platforms with limited computational power. We believe that the proposed tracker can better fit for tracking from such platforms as it is highly efficient and scale adaptive. Taking into account these reasons, it is more appropriate to test it on aerial videos captured by UAVs.

We compare the proposed tracker to other state-of-the-art high speed trackers using the \textit{Precision} and \textit{Success} metrics. Precision is computed as the eucl%

%---------------------------------------------------------------------- 
\section{Conclusion} \label{sc:Conclusion}
%----------------------------------------------------------------------

%--------------
%\section*{Acknowledgements}
%put stuff here for the accepted , but not the ICCV version

\small
\bibliographystyle{ieee}
\bibliography{draft}

\end{document}